Predicting Post-Election FAANG Stock Prices: Accurate November Trends and Flat December Projections

### 2024-12-04 paper/paper.qmd
commit changes to preamble in R scripts

The performance of FAANG stocks - **Meta (META)**, **Amazon (AMZN)**, **Apple (AAPL)**, **Netflix (NFLX)**, and **Google (GOOGL)**, has become a cornerstone of global financial markets, reflecting broader economic trends and investor sentiment. These companies represent diverse sectors, from e-commerce and technology to entertainment and digital advertising, making their stock price movements critical indicators of market health. In the context of the 2024 post-pandemic recovery and a pivotal United States Presidential Election, understanding and predicting FAANG stock price trends is particularly relevant for investors and policymakers alike.

This paper focuses on historical stock price data from years 2020 to 2024, sourced from Yahoo Finance [@yahoofinance] to develop predictive models for FAANG stock prices in November and December 2024. Utilizing XGBoost regression models, we analyze historical trends and engineered features, including lagged prices, moving averages, and volatility measures, to forecast stock price movements. These predictions provide valuable insights into the behavior of FAANG stocks during critical periods, such as post-election market adjustments, and highlight opportunities and risks for short-term and long-term investment strategies.

@@

The implications of these findings extend beyond price predictions, offering a framework to evaluate market behavior in response to external factors such as elections and economic conditions. The remainder of this paper is organized as follows: Section 2 describes the data sources, variables, and preprocessing steps; Section 3 outlines the modeling approach and feature engineering strategies; Section 4 presents the results and discusses trends in predicted stock prices; and Section 5 addresses the implications and limitations of the findings, with additional model diagnostics detailed in the [Appendix -@sec-model-details].

@@

ggplot(dec_results, aes(x = as.Date(date), y = predicted_adjusted, color = symbol)) +
  geom_line(size = 1) +
  labs(
    x = "Date",
    y = "Predicted Stock Price",
    color = "Stock Symbol"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    text = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
@fig-dec-pred reveals flat stock price trends for all FAANG companies, with no significant fluctuations throughout the month. Netflix (NFLX) is projected to maintain the highest price, exceeding $875, while Google (GOOGL) consistently exhibits the lowest price, just above $200. Apple (AAPL), Amazon (AMZN), and Meta (META) display similar stability, with their prices clustered between $250 and $600. These flat trends suggest that the model struggles to capture potential market dynamics and daily volatilities, likely due to the absence of real-world influences such as macroeconomic shifts or investor sentiment in the input data. Further discussion on these limitations is provided in [Appendix -@sec-model-limitations].

@@

## Model Justification - Further Details

@@

>>>>

The performance of FAANG stocks - **Meta (META)**, **Amazon (AMZN)**, **Apple (AAPL)**, **Netflix (NFLX)**, and **Google (GOOGL)**, has become a cornerstone of global financial markets, reflecting broader economic trends and investor sentiment. These companies represent diverse sectors, from e-commerce and technology to entertainment and digital advertising, making their stock price movements critical indicators of market health. As the world navigates the post-pandemic recovery and the uncertainties surrounding the 2024 United States Presidential Election, understanding and forecasting FAANG stock prices is both timely and consequential for investors and policymakers alike.

This paper examines FAANG stock price trends using historical data spanning 2020 to 2024, sourced from Yahoo Finance [@yahoofinance], to develop predictive models for November and December 2024. Utilizing XGBoost regression, the analysis leverages engineered features, including lagged prices, moving averages, and volatility measures, to predict adjusted closing prices for these stocks during the study period. By capturing trends and identifying potential risks and opportunities, these models provide actionable insights for short- and long-term investment strategies.

@@

The implications of these findings extend beyond price predictions, offering a framework to evaluate market behavior in response to external factors such as elections and economic conditions. By contextualizing these trends within the broader market environment, this analysis contributes to a deeper understanding of market dynamics during critical periods. The remainder of this paper is organized as follows: @sec-data describes the data sources, variables, and preprocessing steps; @sec-model outlines the modeling approach and feature engineering strategies; @sec-results presents the results and discusses trends in predicted stock prices; and @sec-discussion addresses the implications and limitations of the findings, with additional model diagnostics detailed in the [Appendix -@sec-model-details].

@@

ggplot(dec_results, aes(x = as.Date(date), y = predicted_adjusted, color = symbol)) +
  geom_line(size = 1) +
  labs(
    x = "Date",
    y = "Predicted Stock Price",
    color = "Stock Symbol"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    text = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
@fig-dec-pred reveals flat stock price trends for all FAANG companies, with no significant fluctuations throughout the month. Netflix (NFLX) is projected to maintain the highest price, exceeding $875, while Google (GOOGL) consistently exhibits the lowest price, just above $200. Apple (AAPL), Amazon (AMZN), and Meta (META) display similar stability, with their prices clustered between $250 and $600. These flat trends suggest that the model struggles to capture potential market dynamics and daily volatilities, likely due to the absence of real-world influences such as macroeconomic shifts or investor sentiment in the input data. Further discussion on these limitations is provided in [Appendix -@sec-model-perform-limit].

@@

## Model Justification - Further Details {#sec-model-justification}

@@

## Model Performance and Limitations {#sec-model-perform-limit}
